---
title: "eds221-0824"
format: html
editor: visual
---

```{r}
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(dplyr)
library(ggplot2)
library(reprex)

```

# Warm-up:

Starting from penguins, create a new subset in which you:

1\) only keep columns species, island, sex, body_mass_g, and flipper_length_mm

2\) Filter to only include Chinstrap and Gentoos

3\) rename the species column to penguin_species

4\) Create a summary table that contains the mean of body mass, the mean of flipper length, and the total count of observations each by species and sex

```{r}
penguins2 <- penguins %>% 
  select(penguin_species = species, island, sex, body_mass_g, flipper_length_mm) %>% filter(penguin_species %in% c("Chinstrap", "Gentoo")) %>% group_by(penguin_species, sex) %>% 
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE), mean_flip_length = mean(flipper_length_mm, na.rm = TRUE), count_obs = n())
```

A few variations on dplyr::across()

Create a summary table of mean values by species for all numeric variables in the penguins dataset

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), max, na.rm = TRUE))


```

calculate the mean value of all columns that start with "bill"

```{r}

penguins %>% 
  group_by(species, island) %>% 
  summarize(across(starts_with("bill"), max, na.rm = TRUE))



```

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(across(ends_with("mm"), min, na.rm = TRUE))

```

find and return both the mean and standard deviation of all columns starting with "bill", grouped by penguin species, and returning a new name that combines the original column with the function

```{r}
  penguins %>% 
  group_by(species) %>% 
summarize(across(starts_with("bill"), list(mean = mean, sd = sd), .names = "{.col}.hello.{.fn}"))

```

```{r}
penguins_table <- 
  penguins %>% 
  group_by(species, sex) %>% 
  sumarize(mean_mass = mean(body_mass_g, na.rm = TRUE), 
           sd_mass = sd(body_mass_g, na.rm = TRUE)) %>% 
  kable(col.names = c("Species"))
```

```{r}
chart1 <- penguins %>%  
  ggplot(aes(x = body_mass_g, y = flipper_length_mm))

chart2<- penguins %>% 
  ggplot(aes(x=bill_length_mm)) + geom_histogram()

```

```{r}
((chart1 + chart2)) / (chart2 + chart1)
```

```{r}
ggplot(penguins, aes(x= species, y = body_mass_g)) + geom_jitter(width = 0.2, alpha = 0.5, color = "navy")
```

```{r}

#| fig-cap: Penguin body mas observations by species. 
#| fig-alt: Jitterplot showing body mass measurements (y-axis, in grams) for three penguin species (Adelie, Chinstrap and Gentoo on x-axis) observed near Palmer Archipelago, Antarctica. Adelie and Chinstraps ahave similar body mass distribution, with Gentoos noticeably larger than both. 

ggplot(penguins, aes(x= species, y = body_mass_g ))
```

```{r}
library(tidyverse)
ggplot(data = mtcars, aes(x= mpg, y = hp)) + geom_point()
```
